From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/finite.tex
Remove word problem

% High-level explanation of technical devices
First, we will define the required notions of free groups and group presentations, and state some of their most
important properties. Then, we introduce our chosen \emph{Coxeter presentation} for $\Sn$. To solve the word problem for
$\Sn$, we will use a rewriting system, with a suitable, well behaved collection of reduction rules corresponding to the
Coxeter presentation. Finally, we describe the normal forms in this rewriting system, using Lehmer codes, and prove the
correspondence between them and the type $\Aut[\Fin[n]]$ of automorphisms on a finite set. The generators and relations
we use here will be used to quote back to 1 and 2-combinators in $\PiLang$ (see~\cref{sec:equivalence}).

@@

We use a presentation based on adjacent transpositions for generators, called a~\emph{Coxeter presentation}, which we
also use to solve the word problem for $\Sn$. Connecting it back to the language $\PiLang$, we will show that it is
possible to encode all $\PiLang$ combinators using adjacent transpositions (in~\Cref{sec:equivalence}). The rest of this
section develops these intuitions formally. Group presentations are built by adding equations to a \emph{free group},
which we define first.

@@

Before, the only way to decide the equality of two elements in a group was to evaluate and check them on the nose, but
in a group presentation, this is reduced to a \emph{word problem}, that is, deciding whether one word -- a
representative of the equivalence class of the group's elements, can be reduced to another word, using the group's
relations. However, these equations are not directed, so it is not always possible to construct a well-behaved rewriting
system. In general, the word problem for groups is known to be undecidable.

@@

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item $\normf$ splits into a section-retraction pair, that is, we have ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n]
            \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$ and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \), where $\quotinc$ is the inclusion into the quotient.
  \end{enumerate}
\end{proposition}

@@

We set out to solve the word problem for $\Sn$. To decide if two words in $\List[\Fin[n]]$ are $\cox*$-equal, we simply
have to compute their normal forms using $\normf$. They correspond to the same permutation if and only if these normal
forms are equal, which is decidable for $\List[\Fin[n]]$.

>>>>

% High-level explanation of technical devices
First, we will define the required notions of free groups and group presentations, and state some of their most
important properties. Then, we introduce our chosen \emph{Coxeter presentation} for $\Sn$. To solve the word problem for
$\Sn$, we will use a rewriting system, with a suitable, well behaved collection of reduction rules corresponding to the
Coxeter presentation equations. Finally, we describe the normal forms in this rewriting system, using Lehmer codes, and prove the
correspondence between them and the type $\Aut[\Fin[n]]$ of automorphisms on a finite set. The generators and relations
we use here will be used to quote back to 1 and 2-combinators in $\PiLang$ (see~\cref{sec:equivalence}).

@@

We will show that it is possible to encode all $\PiLang$ combinators using adjacent transpositions
(in~\Cref{sec:equivalence}). Group presentations are built by adding equations to a \emph{free group}, which we define
first.

@@

Before, the only way to decide the equality of two elements in a group was to evaluate and check them on the nose, but
in a group presentation, this is reduced to deciding whether one word -- a representative of the equivalence class of
the group's elements, can be reduced to another word, using the group's relations. However, these equations are not
directed, so it is not always possible to construct a well-behaved rewriting system. In general, the word problem for
groups is known to be undecidable.

@@

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item $\normf$ splits into a section-retraction pair, that is, we have ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n]
            \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$ and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \), where $\quotinc$ is the inclusion into the quotient, and
    $\im{f}$ denotes the image of $f$ (see~\cref{def:im}).
  \end{enumerate}
\end{proposition}

@@

To decide if two words in $\List[\Fin[n]]$ are $\cox*$-equal, we simply have to compute their normal forms using
$\normf$. They correspond to the same permutation if and only if these normal forms are equal, which is decidable for
$\List[\Fin[n]]$.
